visualizemi: Visualization, Effect Size, and Replication of Measurement Invariance for Registered Reports

### 2023-10-16 manuscript.tex
all the updates

  \usepackage{unicode-math}

@@

@@

\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}

@@

\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}

@@

\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}

@@

\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}

@@

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Visualizing and Interpreting Multi-Group Confirmatory Factor Analysis},
  pdfauthor={Erin M. Buchanan1},
  pdflang={en-EN},
  pdfkeywords={multigroup confirmatory factor analysis, measurement invariance, visualization, effect size},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Visualizing and Interpreting Multi-Group Confirmatory Factor Analysis}
\author{Erin M. Buchanan\textsuperscript{1}}
\date{}

@@

\abstract{%
Latent variable modeling as a lens for psychometric theory is a popular tool for social scientists to examine measurement of constructs (Beaujean, 2014). Journals such as \emph{Assessment} regularly publish articles supporting measures of latent constructs wherein a measurement model is established. Confirmatory factor analysis can be used to investigate the replicability and generalizability of the measurement model in new samples, while multi-group confirmatory factor analysis is used to examine the measurement model across groups within samples (Brown, 2015). With the rise of the replication crisis and ``psychology's renaissance'' (Nelson et al., 2018), interest in divergence in measurement has increased, often focused on small parameter differences within the latent model. This manuscript outlines ways to visualize potential non-invariance, to supplement large numbers of tables that often overwhelm a reader within these published reports. Readers will learn how to interpret the impact and size of the proposed non-invariance in models. While it is tempting to suggest that problems with replication and generalizability are simply issues with measurement, it is crucial to remember that all models have variability and error, even those models estimating the differences between item functioning, such as multi-group confirmatory factor analysis.
}

@@

Measurement invariance is typically analyzed using confirmatory factor analysis, specifically, multi-group confirmatory factor analysis (MGCFA) or less often, with item response theory (Stark et al., 2006; Tay et al., 2015). First, the model is examined with the factor structure proposed for the latent and observed variables, and then often these models are assessed for each group separately. The two models are then combined together into one nested CFA in order to determine configural invariance (Brown, 2015; Byrne, 2001; Kline, 2016). Configural invariance tests if the proposed factor structure is the same between groups. In this model, all other estimated parameters are allowed to vary between groups. The general approach is to use this model as a baseline for starting a sequential analysis of further restrictions between group parameters (i.e., more restrictive with each step). However, models without configural invariance can occur and often point to misspecification for the observed and latent variables within one group (i.e., cross loadings of items onto other latent variables or correlated error terms for one group only).

Next, the estimated parameter between each observed variable and its latent variable are constrained to be equal between groups for metric invariance. For example, item 1's factor loading must be equal to item 1's factor loading for each group. This test examines if the items represent the same relationship to the latent variable, or if specific items have weaker or stronger relationships in specific groups. Finding non-invariance at this stage generally points to items that have different functioning or interpretation for one group. At the third model, the item intercepts (i.e., item averages) are restricted across groups for scalar invariance. Scalar non-invariance would indicate that items have the same strength of relationship with their latent variable, just one group has a higher overall average on that item. Last (although sometimes not used), we may consider constraining error variances for each observed variable to be equal across groups for strict invariance. Strict non-invariance can occur when one group has a higher range of values on the observed variable, thus showing a larger variance. For example, if using a Likert scale, one group may use the full 1 to 7 range (creating a flatter distribution and larger variance), while the other group shows a ceiling effect of only using 5 to 7.

These concepts have been explored and implemented for the last fifty years (Jöreskog, 1971; Sörbom, 1978) and implemented in the most popular structural equation modeling programs (Boker et al., 2011; Jöreskog \& Sörbom, 2001; Rosseel, 2012). Byrne et al. (1989) extended the ideas of multi-group testing by suggesting partial invariance (followed by Meredith, 1993). Partial invariance occurs when non-invariance is found but can be attributed to only a few parameter estimate differences between groups (i.e., items 1 and 2 have different factor loadings but all others are the same). This testing provided an advantage to understand where the potential non-invariance may occur for further study and interpretation guidelines. To determine when non-invariance and partial invariance occurred, each model is sequentially compared to the previous model using some form of a difference test. Traditionally, since models were nested, a chi-square difference test was used (Cheung \& Rensvold, 2002; Meade et al., 2008); however, given the known issues with chi-square (Thompson \& Daniel, 1996), people have favored empirical cutoffs for differences in fit indices. As the field pushes back against favoring cutoff criteria and rules of thumb (Marsh et al., 2004; Putnick \& Bornstein, 2016), an effect size measure for translating ``how much'' non-invariance was developed \(d_{MACS}\) (Nye \& Drasgow, 2011). This effect size examines the differences in observed variables between the two groups for both the factor loading and the item intercept; thus, any differences in either or both will increase the effect size for non-invariance (Stark et al., 2006).

With \(d_{MACS}\) and measurement invariance testing, researchers can begin to quantify how and where their construct measurement may vary between groups. However, given the large number of studies that show non-invariance, it is clear that equivalence can be hard to meet. It is difficult to know if non-invariance occurs because of random sampling error, true population differences, or differences in replication and reproducibility of the construct in a new sample. Further, it is important to remember that the parameter estimates that we are testing are just that - estimates. All the parameter estimates have measures of standard error to indicate that they are more than likely variable with a new sample or population. Given that this information is generally ignored during the examination of measurement invariance, it may be that we are claiming that many scales are non-invariant, when in reality, the differences between loadings or item intercepts are small and unimportant. \(d_{MACS}\) provides the opportunity to begin to think about the smallest effect size of interest or the smallest meaningful effect size (Anvari \& Lakens, 2021; Lakens, 2017). As mentioned, \(d_{MACS}\) has only really been explored for a combined intercept and loadings, and while useful, does not necessarily allow a researcher to pinpoint specific issues within an observed variable. The purpose of this manuscript is provide readers with a framework for visualization of differences in loadings, intercepts, and variances for each item, and the impact of those differences on the distribution of the latent mean. No known visualization techniques have been proposed for measurement invariance. By creating panel visualizations, we can supplement a researchers ability to judge the strength of the non-invariance differences and effect size for each item. Coupled with other indicators (i.e., fit indices differences, \(d_{MACS}\)), we can move toward a better understanding of how much measurement non-invariance is meaningful.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Be able to create visualizations for common steps to multi-group confirmatory factor analysis.
\item
  Be able to interpret the impact and size of potential non-invariance on measurement.
\item
  Understand the impact of measurement variability on replication and generalizability.
\end{enumerate}

@@

Data was simulated using the \texttt{simulateData} function in the \emph{R} package \emph{lavaan} (Rosseel, 2012) assuming multivariate normality using a \(\mu\) of 0 and \(\sigma\) of 1 for the data. This function allows you to write \emph{lavaan} syntax for your model with estimated values to generate data for observed variables. The data included two groups of individuals (``Group 1'', ``Group 2'') for a multi-group confirmatory factor analysis (\(n_{group}\) = 250, \emph{N} = 500). The latent variables were assumed to be continuous normal. The model consisted of five observed items predicted by one latent variable (\texttt{lv\ =\textasciitilde{}\ q1\ +\ q2\ +\ q3\ +\ q4\ +\ q5}); however, the demonstration in this manuscript extends to multiple latent variables and other combinations of observed variables. Each item was assumed to be related to the latent variable with loadings approximately equal to .40 to .80, except when cases of non-invariance on the loadings was assumed.

The Brown (2015) steps of testing measurement invariance are demonstrated in this manuscript for illustration purposes, but in line with Stark et al. (2006) suggestions, the visualizations show the impact of loadings and intercepts together. The configural model was analyzed nesting both groups into the same CFA model requiring that both groups show the same model structure, but all other parameters are free to vary between groups. The metric model constrained the factor loadings of each group to be equal within the model. The scalar model then constrained the item intercepts (i.e., item means) to be equal across groups. Finally, the strict model constrained the item variances (i.e., error variances) to be equal for each item across groups. These models are normally tested sequentially, and a convenience function \texttt{mgcfa} is provided in the supplemental documents for this manuscript. Fit indices for the steps for multi-group models are presented in the appendix for comparison of cutoff rules of thumb (Cheung \& Rensvold, 2002) to effect sizes and visualizations presented in this manuscript. Fit indices include Akaike Information Criterion (AIC, Akaike, 1998), Bayesian Information Criterion (BIC, Schwarz, 1978), Comparative Fit Index (CFI, Bentler, 1990), Tucker Lewis Index (TLI, Tucker \& Lewis, 1973), root mean squared error of approximation RMSEA (Steiger, 1990), and standardized root mean square residual (SRMR, Bentler, 1995).

The data was then simulated to represent invariance across all model steps, small, medium, and large invariance using \(d_{MACS}\) estimated sizes from Nye et al. (2019). While \(d_{MACS}\) is used primarily for an effect size of the (non)-invariance for intercepts and loadings together, a similar approach was taken for the estimation of small, medium, and large effects on the residuals. The effect size is presented for all models, calculated from the \emph{dmacs} package (Dueber, 2023; Nye \& Drasgow, 2011). Only one item in each model was manipulated from the invariant model to create the non-invariant models.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{code-examples}{%
\subsection{Code Examples}\label{code-examples}}

The complete code for this manuscript can be found at \url{https://osf.io/wev5f/}, and the function code for the convenience function for multi-group models and plots is found in the appendix. This tutorial was registered at \url{https://osf.io/vwf4d}, and the example provided at the end of the manuscript was added after that registration. First, we would create our model code in \emph{lavaan} syntax (Rosseel, 2012). The \texttt{lv} latent variable predicts the five measured variables, which are present as columns in our \texttt{df.invariant} data set. You would include the dataframe in the \texttt{data} argument of our function, the name of the grouping variable in quotes for \texttt{group}, and the \emph{lavaan} model syntax in the \texttt{model} argument. The \texttt{mgcfa} function code runs an overall model with all data, regardless of group, each group separately on the model, then the steps described above: configural, metric, scalar, and strict invariance.

\emph{lavaan} automatically sets the mean (i.e., the intercept) for latent variables to zero. If we wish to visualize the impact of the changes in parameter estimates across groups on the latent means, we need to allow the latent mean estimation with \texttt{lv\ \textasciitilde{}\ 1}. However, adding this estimation into our model will create a non-identified model. To solve this problem, you can set one of the intercepts of another variable to a value to scale the model. Here we will set the scale of the model by using \texttt{q1\ \textasciitilde{}\ 0*1}, thus, scaling the expected means to zero. With simulation, this step is easy to know which variable to pick - we set the intercept on the variable we know did not show differences. In real data, you may wish to run the model steps \emph{without} setting this option, examine the results of a configural or separate models, and then add the option for the values most similar. Additionally, you could complete partial invariance steps to determine which value appears most consistent to fix.

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# run our mgcfa function to run all models}
\NormalTok{results.invariant }\OtherTok{\textless{}{-}} \FunctionTok{mgcfa}\NormalTok{(}\AttributeTok{data =}\NormalTok{ df.invariant, }\CommentTok{\#dataframe}
                           \AttributeTok{group =} \StringTok{"group"}\NormalTok{, }
                           \AttributeTok{model =}\NormalTok{ model.overall)}
\CommentTok{\# what is saved for you}
\FunctionTok{names}\NormalTok{(results.invariant)}
\end{Highlighting}
\end{Shaded}

@@

\begin{verbatim}
## [1] "model_coef"       "model_fit"        "model.overall"    "model.group1"    
## [5] "model.group2"     "model.configural" "model.metric"     "model.scalar"    
## [9] "model.strict"
\end{verbatim}

The results are saved as a list and include the following:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  \texttt{model\_coef}: a tidy dataframe with \emph{all} model's coefficients saved from the \emph{lavaan} outputs. Note that we can see that the intercept \texttt{\textasciitilde{}1} is set for question 1 but freely estimated for the latent variable.
\end{enumerate}

@@

\begin{verbatim}
## # A tibble: 10 x 13
##    term     op    estimate std.error statistic   p.value  std.lv std.all std.nox
##    <chr>    <chr>    <dbl>     <dbl>     <dbl>     <dbl>   <dbl>   <dbl>   <dbl>
##  1 "lv =~ ~ =~      1         0         NA     NA         0.780   0.598   0.598 
##  2 "lv =~ ~ =~      0.564     0.0864     6.52   6.99e-11  0.440   0.435   0.435 
##  3 "lv =~ ~ =~      0.748     0.105      7.12   1.09e-12  0.583   0.505   0.505 
##  4 "lv =~ ~ =~      0.338     0.0804     4.20   2.62e- 5  0.264   0.250   0.250 
##  5 "lv =~ ~ =~      0.904     0.120      7.52   5.48e-14  0.705   0.613   0.613 
##  6 "q1 ~1 " ~1      0         0         NA     NA         0       0       0     
##  7 "lv ~1 " ~1     -0.0187    0.0584    -0.320  7.49e- 1 -0.0239 -0.0239 -0.0239
##  8 "q1 ~~ ~ ~~      1.09      0.103     10.6    0         1.09    0.643   0.643 
##  9 "q2 ~~ ~ ~~      0.828     0.0604    13.7    0         0.828   0.811   0.811 
## 10 "q3 ~~ ~ ~~      0.997     0.0786    12.7    0         0.997   0.745   0.745 
## # i 4 more variables: model <chr>, block <int>, group <int>, label <chr>
\end{verbatim}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  \texttt{model\_fit}: a tidy dataframe with \emph{all} model's fit indices saved from the \emph{lavaan} outputs.
\end{enumerate}

@@

\begin{verbatim}
## # A tibble: 6 x 18
##    agfi   AIC   BIC   cfi chisq  npar  rmsea rmsea.conf.high   srmr   tli
##   <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>  <dbl>           <dbl>  <dbl> <dbl>
## 1 0.979 7516. 7579. 0.994  6.37    15 0.0234          0.0697 0.0211 0.988
## 2 0.948 3766. 3819. 0.976  7.79    15 0.0473          0.108  0.0312 0.953
## 3 0.952 3762. 3815. 0.980  7.25    15 0.0424          0.104  0.0322 0.960
## 4 0.950 7528. 7654. 0.978 15.0     30 0.0449          0.0886 0.0317 0.956
## 5 0.942 7529. 7639. 0.954 24.7     26 0.0554          0.0905 0.0476 0.934
## 6 0.952 7523. 7616. 0.964 26.2     22 0.0428          0.0760 0.0488 0.960
## # i 8 more variables: converged <lgl>, estimator <chr>, ngroups <int>,
## #   missing_method <chr>, nobs <int>, norig <int>, nexcluded <int>, model <chr>
\end{verbatim}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
  Saved \emph{lavaan} fitted objects that you can use the \texttt{summary()}, \texttt{parameterEstimates()}, \texttt{fitIndices()}, etc. on. Overall model indicates the model without grouping variables testing all data on the proposed model structure. This model is then tested separately for each group (\texttt{model.group1}, \texttt{model.group2}). The final models follow the Brown (2015) naming convention for sequential steps for testing MGCFA for measurement invariance (\texttt{model.configural}, \texttt{model.metric}, \texttt{model.scalar}, \texttt{model.strict}).
\end{enumerate}

The results from the \texttt{model\_coef} table can then be used directly in the suggested plotting function. The plot outputs will be described below. First, here are the arguments for the function:

>>>>

  \usepackage{unicode-math} % this also loads fontspec

@@

\ifPDFTeX\else
  % xetex/luatex font selection

@@

\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}

@@

\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}

@@

\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}

@@

\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}

@@

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={visualizemi: Visualization, Effect Size, and Replication of Measurement Invariance for Registered Reports},
  pdfauthor={Erin M. Buchanan1},
  pdflang={en-EN},
  pdfkeywords={multigroup confirmatory factor analysis, measurement invariance, visualization, effect size},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{visualizemi: Visualization, Effect Size, and Replication of Measurement Invariance for Registered Reports}
\author{Erin M. Buchanan\textsuperscript{1}}
\date{}

@@

\abstract{%
Latent variable modeling as a lens for psychometric theory is a popular tool for social scientists to examine measurement of constructs (Beaujean, 2014). Journals such as \emph{Assessment} regularly publish articles supporting measures of latent constructs wherein a measurement model is established. Confirmatory factor analysis can be used to investigate the replicability and generalizability of the measurement model in new samples, while multi-group confirmatory factor analysis is used to examine the measurement model across groups within samples (Brown, 2015). With the rise of the replication crisis and ``psychology's renaissance'' (Nelson et al., 2018), interest in divergence in measurement has increased, often focused on small parameter differences within the latent model. This manuscript presents \texttt{visualizemi}, an \emph{R} package that provides functionality to calculate multigroup models, partial invariance, visualizations for (non)-invariance, effect sizes for models and parameters, and potential replication rates compared to random models. Readers will learn how to interpret the impact and size of the proposed non-invariance in models with a focus on potential replication and how to plan for registered reports.
}

@@

Measurement invariance has been explored and implemented for the last fifty years (Jöreskog, 1971; Sörbom, 1978) and implemented in the most popular structural equation modeling programs (Boker et al., 2011; Jöreskog \& Sörbom, 2001; Rosseel, 2012). Byrne et al. (1989) extended the ideas of multi-group testing by suggesting partial invariance (followed by Meredith, 1993). Partial invariance occurs when non-invariance is found but can be attributed to only a few parameter estimate differences between groups (i.e., items 1 and 2 have different factor loadings but all others are the same). This testing provided an advantage to understand where the potential non-invariance may occur for further study and interpretation guidelines. As the field pushes back against favoring cutoff criteria and rules of thumb (Marsh et al., 2004; Putnick \& Bornstein, 2016), an effect size measure for translating ``how much'' non-invariance was developed \(d_{MACS}\) (Nye \& Drasgow, 2011). This effect size examines the differences in observed variables between the two groups for both the factor loading and the item intercept; thus, any differences in either or both will increase the effect size for non-invariance (Stark et al., 2006).

With \(d_{MACS}\) and measurement invariance testing, researchers can begin to quantify how and where their construct measurement may vary between groups. Yet, given the large number of studies that show non-invariance, it is clear that equivalence can be hard to meet. It is difficult to know if non-invariance occurs because of random sampling error, true population differences, or differences in replication and reproducibility of the construct in a new sample. The field of psychology is increasingly interested in pre-registration (i.e., registering plans for a study before data collection, Nosek et al., 2018) and the promotion of transparency in study design, implementation, and analysis (Mayo-Wilson et al., 2021), in addition to supporting replication studies (Zwaan et al., 2018). Registered (replication) reports provide an advantageous avenue for the pre-registration of measurement tests, as they allow a researcher the ability to have their study accepted in principle, regardless of the results of a test of construct validity, reliability, or measurement invariance (Hobson, 2019; Nosek \& Lakens, 2014). However, there are few tools that can provide effect size measures for models, individual parameters, or visualization for researchers to plan for future studies. \(d_{MACS}\) provides the opportunity to begin to think about the smallest effect size of interest or the smallest meaningful effect size for measurement invariance and replication (Anvari \& Lakens, 2021; i.e., two studies with overlapping confidence intervals ``replicate,'' even if the test of measurement invariance does not, Lakens, 2017). As mentioned, \(d_{MACS}\) has only really been explored for a combined intercept and loadings, and while useful, does not necessarily allow a researcher to pinpoint specific issues within an observed variable.

Therefore, purpose of this manuscript is describe an \emph{R} package, \texttt{visualizemi}, that provides functionality to calculate multi-group confirmatory factor analysis, partial invariance tests, visualizations of the size of non-invariance, and potential effect sizes for overall models and individual parameters. No known visualization techniques have been proposed for measurement invariance. By creating panel visualizations, we can supplement a researchers ability to judge the strength of the non-invariance differences and effect size for each item. The proposed effect sizes demonstrate the likelihood of replication with a similar sample as compared to a randomly assigned group model, thus, illustrating what type of measurement one might expect to find, and how different that is from random chance. Within this technique, the individual parameter effect sizes can calculated: both the group differences within a model as compared to random and the likelihood of a parameter replication compared to random groups. Coupled with other indicators (i.e., fit indices differences, \(d_{MACS}\)), we can move toward a better understanding of how much measurement non-invariance is meaningful. This tutorial and package will help researchers plan future studies and aid in the ability to estimate a smallest effect of interest for measurement invariance studies, rather than relying on fit indices and rules of thumb alone.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Learn how to use \emph{visualizemi} to analyze multi-group confirmatory factor analysis, examine partial invariance, and create visualizations of parameters.
\item
  Learn how to estimate the potential replication of multi-group models and their parameters using bootstrapping compared to a random group model.
\item
  Be able to calculate and interpret effect sizes for model and parameter replication, as well as parameter group differences.
\item
  Understand the impact of measurement variability on replication and generalizability.
\end{enumerate}

The tutorial will start with simulated data based on known effect sizes using \(d_{MACS}\) and demonstrating the package functions for 1) running the multigroup analysis, 2) running a partial invariance analysis, 3) plotting the partial invariance, 4) estimating replication and effect sizes at the model level, and 5) estimating replication and effect sizes at the parameter level. Last, data from Aiena et al. (2014) examining the measurement invariance of the RS-14 (Wagnild, 2009) will be used to demonstrate the application of the package on real data. The \emph{visualizemi} package vignette includes an additional tutorial walk through.

@@

Data was simulated using the \texttt{simulateData} function in the \emph{R} package \emph{lavaan} (Rosseel, 2012) assuming multivariate normality using a \(\mu\) of 0 and \(\sigma\) of 1 for the data. This function allows you to write \emph{lavaan} syntax for your model with estimated values to generate data for observed variables (see supplemental for examples). The data included two groups of individuals (``Group 1'', ``Group 2'') for a multi-group confirmatory factor analysis (\(n_{group}\) = 250, \emph{N} = 500). The latent variables were assumed to be continuous normal (the package functions do not require this assumption). The model consisted of five observed items predicted by one latent variable (\texttt{lv\ =\textasciitilde{}\ q1\ +\ q2\ +\ q3\ +\ q4\ +\ q5}); however, the demonstration in this manuscript extends to multiple latent variables and other combinations of observed variables. Each item was assumed to be related to the latent variable with loadings approximately equal to .40 to .80, except when cases of non-invariance on the loadings was simulated.

The Brown (2015) steps of testing measurement invariance are demonstrated in this manuscript for illustration purposes, but in line with Stark et al. (2006) suggestions, the visualizations show the impact of loadings and intercepts together. A convenience function \texttt{mgcfa} is used for these steps or other measurement invariance test orders and combinations. Fit indices for the steps for multi-group models are presented in the appendix for comparison of cutoff rules of thumb (Cheung \& Rensvold, 2002) to effect sizes and visualizations presented in this manuscript. Fit indices include Akaike Information Criterion (AIC, Akaike, 1998), Bayesian Information Criterion (BIC, Schwarz, 1978), Comparative Fit Index (CFI, Bentler, 1990), Tucker Lewis Index (TLI, Tucker \& Lewis, 1973), root mean squared error of approximation RMSEA (Steiger, 1990), and standardized root mean square residual (SRMR, Bentler, 1995).

The data was then simulated to represent invariance across all model steps, small, medium, and large invariance using \(d_{MACS}\) estimated sizes from Nye et al. (2019). While \(d_{MACS}\) is used primarily for an effect size of the (non)-invariance for intercepts and loadings together, a similar approach was taken for the estimation of small, medium, and large effects on the residuals. The effect size is presented for all models, calculated from the \emph{dmacs} package (Dueber, 2023; Nye \& Drasgow, 2011). Only one item in each model was manipulated from the invariant model to create the non-invariant models. Given the data was simulated with a \emph{z}-score scaling, the loading values were simulated at .30 points apart (given \(d_{MACS}\) suggestions of .2, .4, .7), the intercepts at .25 points apart, and the residuals at .25 points apart. To plan a simulation for your own study, these values can be used to simulate small, medium, and large non-invariance effects by first converting data into \emph{z}-score.

\hypertarget{package-code-examples}{%
\section{Package Code Examples}\label{package-code-examples}}

The complete code for this manuscript can be found at \url{https://osf.io/wev5f/}. This tutorial was registered at \url{https://osf.io/vwf4d}, and the example provided at the end of the manuscript was added after that registration. The \emph{R} package and replication/effect sizes was added after the original manuscript submission.

\hypertarget{multi-group-cfa-caculation}{%
\subsection{Multi-group CFA Caculation}\label{multi-group-cfa-caculation}}

First, we would create our model code in \emph{lavaan} syntax (Rosseel, 2012). The \texttt{lv} latent variable predicts the five measured variables, which are present as columns in our \texttt{df.invariant} data set. The package does generally require raw data for bootstrapping purposes, and an example of how to simulate data from models and covariance/correlations tables sometimes provided in manuscripts (rather than the raw data) is provided in the supplemental documentation.

\emph{lavaan} automatically sets the mean (i.e., the intercept) for latent variables to zero. If we wish to visualize the impact of the changes in parameter estimates across groups on the latent means, we need to allow the latent mean estimation with \texttt{lv\ \textasciitilde{}\ 1}. However, adding this estimation into our model will create a non-identified model. To solve this problem, you can set one of the intercepts of another variable to a value to scale the model. Here we will set the scale of the model by using \texttt{q1\ \textasciitilde{}\ 0*1}, thus, scaling the expected means to zero. With simulation, this step is easy to know which variable to pick - we set the intercept on the variable we know did not show differences. In real data, you may wish to run the model steps \emph{without} setting this option, examine the results of a configural or separate models, and then add the option for the values most similar. Additionally, you could complete partial invariance steps to determine which value appears most consistent to fix the estimate.

@@

The \texttt{mgcfa} function is designed to flexibly allow you to leverage \texttt{lavaan}'s package functions to calculate multiple measurement steps at once. You would include:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  the model syntax in the \texttt{model} argument
\item
  the dataframe in the \texttt{data} argument of our function
\item
  the name of the grouping variable in quotes for \texttt{group}
\item
  and the equality constraints you would like to impose in order in \texttt{group.equal}
\item
  \texttt{...} any other \emph{lavaan} arguments you would like to use such as \texttt{meanstructure} or \texttt{estimator}.
\end{enumerate}

Note: you can also use \texttt{sample.cov}, \texttt{sample.mean}, \texttt{sample.nobs} in this step for estimation of multi-group models, but simulated dataframes are needed for bootstrapping replication estimates.

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# run our mgcfa function to run all models}
\NormalTok{results.invariant }\OtherTok{\textless{}{-}} 
  \CommentTok{\# name of the saved model syntax}
  \FunctionTok{mgcfa}\NormalTok{(}\AttributeTok{model =}\NormalTok{ model.overall, }
        \CommentTok{\# name of the dataframe}
        \AttributeTok{data =}\NormalTok{ df.invariant,}
        \CommentTok{\# name of the grouping variable}
        \AttributeTok{group =} \StringTok{"group"}\NormalTok{,}
        \CommentTok{\# equality constraints to impose in order}
        \AttributeTok{group.equal =} \FunctionTok{c}\NormalTok{(}\StringTok{"loadings"}\NormalTok{, }\StringTok{"intercepts"}\NormalTok{, }\StringTok{"residuals"}\NormalTok{),}
        \CommentTok{\# other options to send to lavaan cfa function}
        \AttributeTok{meanstructure =}\NormalTok{ T)}

\CommentTok{\# what is saved for you}
\FunctionTok{names}\NormalTok{(results.invariant)}
\end{Highlighting}
\end{Shaded}

@@

\begin{verbatim}
## [1] "model_coef"        "model_fit"         "model_overall"    
## [4] "group_models"      "model_configural"  "invariance_models"
\end{verbatim}

The following output is saved:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  \texttt{model\_coef}: The parameter estimates for each model with the model step included in a \emph{model} column. This set of coefficients can be used for other functions. This dataframe is created with \emph{broom}'s \texttt{tidy()} function if you wish to recreate this table without running the \texttt{mgcfa()} function (Robinson et al., 2023).
\end{enumerate}

@@

\begin{verbatim}
## # A tibble: 10 x 13
##    term     op    estimate std.error statistic   p.value  std.lv std.all std.nox
##    <chr>    <chr>    <dbl>     <dbl>     <dbl>     <dbl>   <dbl>   <dbl>   <dbl>
##  1 "lv =~ ~ =~      1         0         NA     NA         0.803   0.616   0.616 
##  2 "lv =~ ~ =~      0.655     0.0880     7.44   9.77e-14  0.526   0.493   0.493 
##  3 "lv =~ ~ =~      0.640     0.0895     7.15   8.83e-13  0.514   0.463   0.463 
##  4 "lv =~ ~ =~      0.277     0.0749     3.69   2.24e- 4  0.222   0.209   0.209 
##  5 "lv =~ ~ =~      0.955     0.117      8.13   4.44e-16  0.766   0.656   0.656 
##  6 "q1 ~1 " ~1      0         0         NA     NA         0       0       0     
##  7 "lv ~1 " ~1     -0.0305    0.0582    -0.524  6.00e- 1 -0.0380 -0.0380 -0.0380
##  8 "q1 ~~ ~ ~~      1.05      0.0995    10.6    0         1.05    0.620   0.620 
##  9 "q2 ~~ ~ ~~      0.860     0.0653    13.2    0         0.860   0.757   0.757 
## 10 "q3 ~~ ~ ~~      0.966     0.0711    13.6    0         0.966   0.785   0.785 
## # i 4 more variables: model <chr>, block <int>, group <int>, label <chr>
\end{verbatim}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  \texttt{model\_fit}: The model fit indices from \texttt{fitmeasures()} to review for overall model fit and invariance judgments. The name of the model is included in a \emph{model} column.
\end{enumerate}

@@

\begin{verbatim}
## # A tibble: 6 x 18
##    agfi   AIC   BIC   cfi  chisq  npar  rmsea rmsea.conf.high    srmr   tli
##   <dbl> <dbl> <dbl> <dbl>  <dbl> <dbl>  <dbl>           <dbl>   <dbl> <dbl>
## 1 0.998 7516. 7580. 1      0.650    15 0               0      0.00616 1.04 
## 2 0.948 3766. 3819. 0.976  7.79     15 0.0473          0.108  0.0312  0.953
## 3 0.974 3768. 3820. 1      4.48     15 0               0.0831 0.0210  1.01 
## 4 0.961 7533. 7660. 0.991 12.3      30 0.0301          0.0785 0.0261  0.982
## 5 0.965 7528. 7638. 0.994 15.4      26 0.0200          0.0660 0.0330  0.992
## 6 0.969 7522. 7615. 1     17.3      22 0               0.0542 0.0352  1.00 
## # i 8 more variables: converged <lgl>, estimator <chr>, ngroups <int>,
## #   missing_method <chr>, nobs <int>, norig <int>, nexcluded <int>, model <chr>
\end{verbatim}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
  \texttt{model\_overall}: A saved \emph{lavaan} fitted model of all groups together without any equality constraints or grouping variables. These objects can be used with any function that normally takes a saved model: \texttt{parameterEstimates()}, \texttt{modificationIndices()}, \texttt{semPlot::semPaths()}, and so on (Epskamp, 2022).
\item
  \texttt{group\_models}: A list of saved fitted models for each group separately.
\item
  \texttt{model\_configural}: A saved fitted model for the configural model that nests together each group into one model with no other constraints.
\item
  \texttt{invariance\_models}: A list of saved fitted models that consecutively adds \texttt{group.equal} constraints.
\end{enumerate}

\hypertarget{package-function}{%
\subsubsection{Package Function}\label{package-function}}

The results from the \texttt{model\_coef} table can then be used directly in \texttt{plot\_mi()}. The plot outputs will be described below. First, here are the arguments for the function:
